\subsection{Variable templates}

\begin{frame}[t,fragile]{Variable template}
\begin{itemize}
  \item A \textmark{variable template} defines a family of variables.
    \begin{itemize}
      \item One variable is generated for each instantiation.
    \end{itemize}
\end{itemize}

\begin{block}{Counting uses per type}
\begin{lstlisting}
template <typename T>
int use_count = 0;

template <typename T>
void log(T x) {
  // Note: This is not thread-safe
  std::cout << std::format("log: {}, uses = {}\n", x, ++use_count<T>);
}
\end{lstlisting}
\end{block}
\end{frame}

\begin{frame}[t,fragile]{Using variable templates for constants}
\begin{itemize}
  \item Variable templates can be used to define generic compile-time constants.
\end{itemize}

\begin{block}{Generic constants}
\begin{lstlisting}
template <std::floating_point T>
constexpr T pi = T(3.1415926535897932385); 

template <std::floating_point T>
T circle_area(T r) {
  return pi<T> * r * r;
}

template <std::floating_point T>
void print_area(T x) {
  std::cout << std::format("area({}) = {}\n", x, circle_area(x));
}
\end{lstlisting}
\end{block}

\end{frame}


\begin{frame}[t,fragile]{Mathematical constants in the standard library}
\begin{itemize}
  \item The standard library offers common mathematical constants as variable constant templates.
    \begin{itemize}
      \item Header \cppid{<numbers>}.
      \item Namespace \cppid{std::numbers}.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item Generic version with \cppid{\_v} suffix.
    \begin{itemize}
      \item \cppid{e\_v}, \cppid{pi\_v}, \cppid{log2e\_v}, \cppid{log10e\_v}, \cppid{inv\_pi\_v}, \ldots
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item Non-generic \cppkey{double} version without \cppid{\_v} suffix.
    \begin{itemize}
      \item \cppid{e}, \cppid{pi}, \cppid{log2e}, \cppid{log10e}, \cppid{inv\_pi}, \ldots
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Metaprogramming predicates in the standard library}
\begin{itemize}
  \item The standard library offers many predicates used in metaprogramming.
    \begin{itemize}
      \item \cppid{std::is\_integral<T>}, \cppid{is::arithmetic<T>}, \cppid{is::enum<T>}, \ldots
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item All of them have a \cppid{value} member.
\begin{lstlisting}
if constexpr(std::is_arithmetic<T>::value) { /*...*/ }
\end{lstlisting}

  \mode<presentation>{\vfill\pause}
  \item The library offers shortcuts as variable constant templates.
\begin{lstlisting}
if constexpr(std::is_arithmetic_v<T>) { /*...*/ }
\end{lstlisting}
\end{itemize}
\end{frame}
