\subsection{Two phase translation model}

\begin{frame}[t,fragile]{Translation model}
\begin{itemize}
  \item A generic function is translated in two \textmark{stages} or \textmark{phases}.

  \begin{enumerate}

    \mode<presentation>{\vfill\pause}
    \item \textmark{Definition time}:
          The code is checked without considering generic types porperties.
      \begin{itemize}
        \item Check \textbad{unknown names} that do not depend on generic types.
        \item Check \textbad{syntax errors}.
        \item Evaluate \textbad{static assertions} that do not depend on generic types.
      \end{itemize}
    

    \mode<presentation>{\vfill\pause}
    \item \textmark{Instantiation time}.
      \begin{itemize}
        \item Check all parts that \textmark{depend on generic types}.
      \end{itemize}

  \end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Checking at definition time}
\begin{block}{Errors checked at definition time}
\begin{lstlisting}
void print_square(std::integral auto x) {
  print("hello"); // Undeclared identifier print

  std::integral auto r = x * x // Syntax error: missing semicolon

  static_assert(sizeof(int) >= sizeof(long)); // static_assert failed
}
\end{lstlisting}
\end{block}

  \begin{itemize}
    \item \textbad{Unknown functions} can be diagnosed 
          when they do not depend on the type of \cppid{x}.

    \item \textbad{Syntax errors} can be found 
          without knowing the exact type of \cppid{x}.

    \item \textbad{Static assertions} that do not depend 
          on generic properties can be evaluated.
  \end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Checking at instantiation time}

\begin{columns}[T]

\column{.6\textwidth}

\begin{block}{Errors checked at instatiation time}
\begin{lstlisting}
void print_value(int x) { std::cout << x << '\n'; }
void print_value(long) = delete;

void print_square(std::integral auto x) {
  print_value(x); // OK if instantiated with x of type int
  std::integral auto r = x * x;
  static_assert(sizeof(x) == sizeof(int));
}

int main() {
  print_square(2); // OK
  print_square(2L); // Error: cannot instantiate for long
}
\end{lstlisting}
\end{block}

\column{.4\textwidth}

\pause
\begin{itemize}
  \item Instantiation time errors \textmark{depend} on properties 
        \textbad{unknown} until a \textmark{specific instantiation} occurs.

  \mode<presentation>{\vfill\pause}
  \item When instantiation happens the defintion must be visible.

  \mode<presentation>{\vfill\pause}
  \item \textbad{Consequence}: Generic functions need to be 
        defined before they are used.
    \begin{itemize}
      \item Place generic functions in header files.
    \end{itemize}
\end{itemize}

\end{columns}

\end{frame}
