\subsection{Argument Dependent Lookup}

\begin{frame}[t,fragile]{Ordinary lookup}
\begin{itemize}
  \item When a function (or an operator) is invoked ordinary lookup is performed.

  \mode<presentation>{\vfill\pause}
  \item Ordinary lookup may find one of the following:
    \begin{itemize}
      \item The name of a member function.
      \item The name of a variable.
      \item The name of a type.
      \item The name of local function declaration.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item Otherwise ADL is used.
    \begin{itemize}
      \item \textmark{ADL}: Argument Dependent Lookup.
      \item It solves the interaction between 
            \cppkey{template} and \cppkey{namespace}.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textbad{Warning}: \textmark{ADL} is inhibited if the function name is
        enclosed in parentheses.
\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{ADL}
\begin{itemize}
  \item When \textmark{ADL} is applied, lookup is performed in
    \begin{itemize}
      \item Namespaces associated with the types of call arguments.
      \item Classes associated with the types of call arguments.
    \end{itemize}
\end{itemize}

\begin{block}{ADL in practice}
\begin{lstlisting}[basicstyle=\mode<presentation>{\tiny}]
namespace geom {
  struct point { double x, y; };
  void print(point p);
}

template <typename T>
void print_value(T value) {
  print(value);
}

void f() {
  geom::point origin{0.0, 0.0};
  print_value(origin); // Invoke print(point)
}
\end{lstlisting}
\end{block}

\end{frame}

\begin{frame}[t,fragile]{ADL}

\mode<presentation>{\vspace{-1em}}
\begin{columns}[T]

\column{.6\textwidth}
\begin{block}{Enumerations}
\begin{lstlisting}
namespace graphics {
  enum class color : std::int8_t { red, green, blue };
  std::string to_string(color c);
  void log(color c) {
    std::cout << std::format("COLOR: {}\n", to_string(c));
  }
}

void log(int x) {
  std::cout << std::format("INT: {}\n", x);
}

void f() {
  log(42); // Built in: No ADL. Prints INT: 42
  log(graphics::color::red); // ADL: graphics. Prints COLOR: red
}
\end{lstlisting}
\end{block}

\column{.4\textwidth}
\begin{itemize}
  \item Built-in types $\Rightarrow$ Empty set.
  \item Enumeration type $\Rightarrow$ Enclosing namespace/class.
\end{itemize}

\end{columns}
\end{frame}

\begin{frame}[t,fragile]{Other rules for associated set}
\begin{itemize}
  \item Pointer and array types $\Rightarrow$ Associated set of the underlying type.
  \item Class members $\Rightarrow$ Enclosing class.
  \item Class types $\Rightarrow$ Class, enclosing class and direct/indirect base classes.
  \item Class types $\Rightarrow$ Enclosing namespace.
  \item Class template instance $\Rightarrow$ Types of template type arguments, 
        classes/namespaces where arguments are declared.
  \item Function types $\Rightarrow$ namespaces/classes associated with parameter types and return type.
  \item Pointer to member $\Rightarrow$ namespaces/classes associated with class and with type of member.
\end{itemize}
\end{frame}

