\subsection{Using a generic class}

\begin{frame}[t,fragile]{Using a class template}
\begin{itemize}
  \item Using a class template requires providing explicitly concrete template arguments.
    \begin{itemize}
      \item An exception is using CTAD (class template argument deduction).
    \end{itemize}

\begin{block}{Defining objects from generic classes}
\begin{lstlisting}
void f() {
  fixed_vector<double> marks{5};
  marks.push_back(1.5);
  auto new_marks = marks;

  fixed_vector<std::string> names{4};
  names.push_back("Daniel");
  names.push_back("Maria");
  std::cout << names.size() << '\n';
}
\end{lstlisting} 
\end{block}

\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Using class templates and requirements}
\begin{itemize}
  \item The \textmark{concrete template arguments} must satisfy the \textemph{requirements}
        from \textmark{class template parameters}.

\begin{block}{Requirements satisfaction}
\begin{lstlisting}
void f() {
  fixed_vector<std::string> names{5}; // OK
  fixed_vector<std::optional<std::string>> op_names{5}; // OK
  fixed_vector<std::unique_ptr<std::string>> ptr_names{5}; // Error
}
\end{lstlisting}
\end{block}

  \mode<presentation>{\vfill}
  \item \textmark{Reason}: \cppid{std::unique\_ptr} is not \cppid{std::semiregular}.
    \begin{itemize}
      \item It is \textbad{not} \cppid{std::copyable}.
      \item It is \textbad{not} \cppid{std::copy\_constructible}.
    \end{itemize}

\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Generic classes and instantiation}
\begin{itemize}
  \item Instantitation really happens at member function level.
    \begin{itemize}
      \item Only member functions that are invoked are instantiated.
    \end{itemize}

\begin{columns}[T]

\column{.5\textwidth}

\begin{block}{Defining objects from generic classes}
\begin{lstlisting}
void f() {
  fixed_vector<double> marks{5};
  marks.push_back(1.5);
  auto new_marks = marks;

  fixed_vector<std::string> names{4};
  names.push_back("Daniel");
  names.push_back("Maria");
  std::cout << names.size() << '\n';
}
\end{lstlisting} 
\end{block}

\pause
\column{.5\textwidth}

  \item \cppid{fixed\_vector<double>}:
    \begin{itemize}
      \item Constructor from size.
      \item \cppid{push\_back()}.
      \item Copy constructor.
    \end{itemize}

  \pause
  \item \cppid{fixed\_vector<std::string>}:
    \begin{itemize}
      \item Constructor from size.
      \item \cppid{push\_back()}.
      \item \cppid{size()}.
    \end{itemize}

\end{columns}

\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Instantiation requirements}
\mode<presentation>{\vspace{-.75em}}
\begin{itemize}
  \item Instantiation time checks are performed per member function.
    \begin{itemize}
      \item A given member function might be correct only for some instantiations.
    \end{itemize}
\end{itemize}

\mode<presentation>{\vspace{-.75em}}
\begin{columns}[T]

\column{.55\textwidth}

\begin{block}{A serializable \textbf{fixed\_vector}}
\begin{lstlisting}
template <std::semiregular T>
class fixed_vector {
public:
  //...
  void serialize(std::ostream & os) {
    if (size_ <= 0) { return; }
    os << buffer_[0];
    for (int i = 1; i < size_; ++i) {
      os << " , " << buffer_[gsl::narrow<std::size_t>(i)];
    }
  }
  //...
}
\end{lstlisting}
\end{block}

\pause
\column{.45\textwidth}
\begin{block}{Using \textbf{serialize()}}
\begin{lstlisting}
void f() {
  fixed_vector<double> marks{10};
  marks.push_back(1.5);
  marks.push_back(2.5);
  marks.push_back(3.5);
  marks.serialize(std::cout); // OK

  using entry = std::tuple<std::string, double>;
  fixed_vector<entry> students{4};
  students.push_back({"John", 5});
  students.serialize(std::cout); // Error
}
\end{lstlisting}
\end{block}

\end{columns}

\end{frame}
