\subsection{Views over contiguous sequences}

\begin{frame}[t,fragile]{What is a \textbf{std::span}}
\begin{itemize}
  \item A \cppid{span} is a view over a sequence of objects that
        are contiguous in memory.
    \begin{itemize}
      \item As a view, it has reference semantics.
      \item Memory is managed by the source.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item Kinds:
    \begin{itemize}
      \item \textmark{Fixed extent}: The number of elements is fixed
            and known at compile time.
        \begin{itemize}
          \item Internally representable with a pointer.
        \end{itemize}
      \item \textmark{Dynamic extent}: The numbe of elements 
            might change during run-time.
        \begin{itemize}
          \item Internally representable with a pointer and a size
                or with two pointers.
        \end{itemize}
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textbad{Note}: Fixed extent spans might have a smaller representation
        in memory.
\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Fixed extent spans}
\begin{itemize}
  \item A \textgood{fixed extent span} is parametrized by
        its \textmark{element type} and \textmark{compile-time size}.
\begin{lstlisting}
template <typename T, std::size_t Extent>
class span;
\end{lstlisting}

    \mode<presentation>{\vfill\pause}
    \item A \textgood{fixed extent span} can be constructed from
          \textmark{primitive arrays}, \cppid{std::array}s 
          or \textmark{contiguous containers} as \cppid{std::vector}s.
\begin{lstlisting}
  int arr[]{1, 2, 3, 4, 5};
  std::span<int, 5> spn1{arr};

  std::array sarr{1, 3, 5, 7};
  std::span<int,4> spn2{sarr};

  std::vector vec{1, 4, 9};
  std::span<int,3> spn3{vec};
\end{lstlisting}
\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Fixed extent spans}
\begin{itemize}
  \item A \textgood{dynamic extent span} is parametrized by
        its \textmark{element type}.
    \begin{itemize}
      \item Its size is set to \cppid{std::dynamic\_extent}.
    \end{itemize}
\begin{lstlisting}
template <typename T, std::size_t Extent = std::dynamic_extent>
class span;
\end{lstlisting}

    \mode<presentation>{\vfill\pause}
    \item A \textgood{dynamic extent span} can be constructed by
          omitting the extent template argument.
\begin{lstlisting}
  int arr[]{1, 2, 3, 4, 5};
  std::span<int> spn1{arr};

  std::array sarr{1, 3, 5, 7};
  std::span<int> spn2{sarr};

  std::vector vec{1, 4, 9};
  std::span<int> spn3{vec};
\end{lstlisting}
\end{itemize}
\end{frame}
