\subsection{Constraining multiple generic parameters}

\begin{frame}[t,fragile]{A generic maximum function}
\begin{itemize}
  \item Let's consider a \cppid{maximum} function:

\begin{block}{A generic maximum function}
\begin{lstlisting}
arithmetic auto maximum(arithmetic auto x, arithmetic auto y) {
  return (x>y) ? x : y;
}
\end{lstlisting}
\end{block}

  \mode<presentation>{\vfill\pause}
  \item \textmark{Requirements}:
    \begin{itemize}
      \item \cppid{x} and \cppid{y} must be of arithmetic types.
      \item \cppid{x>y} must be a boolean expression.
      \item The result must be of an arithmetic type.
    \end{itemize}

\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Extending maximum() for more types}
\begin{itemize}
  \item Currently \cppid{maximum()} only works for arithmetic types.
    \begin{itemize}
      \item What about other comparable types?
        \begin{itemize}
          \item \cppid{std::string}, \cppid{std::tuple}, \ldots
        \end{itemize}
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textbad{Idea}: Use \cppid{std::totally\_ordered\_with<T,U>} concept.
    \begin{itemize}
      \item Types \cppid{T} and \cppid{U} satisfy the \cppid{std::totally\_ordered\_with} concept
            if:
        \begin{itemize}
          \item All comparison operators for a value \cppid{x} of type \cppid{T} and
                a value \cppid{y} of type \cppid{U} results are consistent 
                with a \textmark{strict total order}.
          \item Comparison operators are:
                \cppkey{==}, \cppkey{!=}, \cppkey{<}, \cppkey{<=}, \cppkey{>}, and \cppkey{>=}.
        \end{itemize}
    \end{itemize}

\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Extended template notation}
\begin{block}{An extended generic function}
\begin{lstlisting}
template <typename T, typename U>
  requires totally_ordered_with<T,U>
auto maximum(T x, U y) {
  return (x>y) ? x : y;
}
\end{lstlisting}
\end{block}

\begin{itemize}
  \mode<presentation>{\vfill}
  \item The \cppkey{requires} clause here is useful for expressing requirements on multiple
        generic parameters.

  \mode<presentation>{\vfill}
  \item Function \cppid{maximum} is defined for any pair of types \cppid{T} and \cppid{U}
        that are totally ordered.
\end{itemize}

\end{frame}

\begin{frame}[t,fragile]{Adding constraints}

\begin{block}{Maximum on strings versus pointers}
\begin{lstlisting}
void f() {
  using namespace std::literals;
  auto r1 = maximum("carlos"s,"daniel"s); // r1 = "daniel"

  auto r2 = maximum("carlos","daniel"); // r2 = "carlos". Ouch!
}  
\end{lstlisting}
\end{block}

\begin{itemize}
  \item When \cppid{maximum} is invoked with two \cppid{std::string} objects,
        operator \cppkey{<} is invoked and lexicographical order is applied.

  \mode<presentation>{\vfill}
  \item When \cppid{maximum} is invoked with two \cppid{char*} objectis,
        operator \cppkey{<} is invoked and memory addresses order is applied.
\end{itemize}

\end{frame}


\begin{frame}[t,fragile]{Constraining types}
\begin{block}{Constraining with static assertions}
\begin{lstlisting}
template<typename T, typename U>
    requires std::totally_ordered_with<T, U>
auto maximum(T x, U y) {
  static_assert(not std::is_pointer_v<T>, "T must not be a pointer type");
  static_assert(not std::is_pointer_v<U>, "T must not be a pointer type");
  return (x > y) ? x : y;
}
\end{lstlisting}
\end{block}

\begin{itemize}
  \item The following constraint must be satisfied:
    \begin{itemize}
      \item Types \cppid{T} and \cppid{U} must define a total order.
    \end{itemize}
  \item The following additional compile-time checks are performed.
    \begin{itemize}
      \item Type \cppid{T} must not be a pointer type.
      \item Type \cppid{U} must not be a pointer type.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Constraining types}
\begin{block}{A more constrained generic function}
\begin{lstlisting}
template<typename T, typename U>
    requires std::totally_ordered_with<T, U>
        and (not std::is_pointer_v<T>)
        and (not std::is_pointer_v<U>)
auto maximum(T x, U y) {
  return (x > y) ? x : y;
}
\end{lstlisting}
\end{block}

\begin{itemize}
  \item Now the following constraints must be satisfied:
    \begin{itemize}
      \item Types \cppid{T} and \cppid{U} must define a total order.
      \item Types \cppid{T} must not be a pointer type.
      \item Types \cppid{U} must not be a pointer type.
    \end{itemize}
\end{itemize}
\end{frame}
