\begin{frame}[t,fragile]{Alternatives for return types in function templates}

\begin{columns}[T]

\column{.5\textwidth}
\begin{itemize}
  \item \textgood{Options}: 
    \begin{itemize}
      \item \textmark{Return by value}: A copy is returned.
        \begin{itemize}
          \item May result in a \textemph{move} or even \textemph{optimized out}.
        \end{itemize}
      \item \textmark{Return by reference}: A reference to an object is returned.
        \begin{itemize}
          \item May result in a dangling reference.
          \item There may be some valid cases: \cppkey{operator[]}, \cppkey{operator<{}<}, \ldots
          \item Returnging read-only through const reference.
        \end{itemize}
    \end{itemize}
\end{itemize}

\pause
\column{.5\textwidth}
\begin{itemize}
  \item In generic functions the value may deduced and explictly specified.
\end{itemize}
\begin{block}{A generic call function}
\begin{lstlisting}
template <typename F, typename T>
T call(F fun, T value) {
  std::cout << std::format("input = {}\t", value);
  T result = fun(value);
  std::cout << std::format("result = {}\t", result);
  return result;
}
\end{lstlisting}
\end{block}

\end{columns}
 
\end{frame}


\begin{frame}[t,fragile]{Dangers with function template return types}
\begin{itemize}
  \item Even if the return type seems a value it might become a reference type.
\end{itemize}

\begin{block}{An increment function}
\begin{lstlisting}
template <typename T>
T increment(T value) {
  std::cout << std::format("input = {}\t", value);
  T result = ++ value;
  std::cout << std::format("result = {}\t", result);
  return result;
}

void f() {
  double x = 3;
  auto r1 = increment(x); // OK
  auto r2 = increment<double&>(x); // Dangling?
}
\end{lstlisting}
\end{block}
\end{frame}

\begin{frame}[t,fragile]{Removing dangers in return type}
\begin{itemize}
  \item Make sure that the return type is not a reference type.
\end{itemize}
\begin{block}{An increment function}
\begin{lstlisting}
template <typename T>
std::remove_reference_t<T> increment(T value) {
  std::cout << std::format("input = {}\t", value);
  auto result = ++ value;
  std::cout << std::format("result = {}\t", result);
  return result;
}

void f() {
  double x = 3;
  auto r1 = increment(x); // OK
  auto r2 = increment<double&>(x); // OK
}
\end{lstlisting}
\end{block}
\end{frame}

\begin{frame}[t,fragile]{Removing dangers in return type with auto}
\begin{itemize}
  \item Specifying \cppkey{auto} as return type triggers return type deduction.
\end{itemize}
\begin{block}{An increment function}
\begin{lstlisting}
template <typename T>
auto increment(T value) {
  std::cout << std::format("input = {}\t", value);
  auto result = ++ value;
  std::cout << std::format("result = {}\t", result);
  return result;
}

void f() {
  double x = 3;
  auto r1 = increment(x); // OK
  auto r2 = increment<double&>(x); // OK
}
\end{lstlisting}
\end{block}
\end{frame}
